Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-02-06 paper.qmd
Updated Figure 12
Added linked plot section

It’s important to access the \gD{} layout and the generated model overlaid on data in \pD{} together to understand whether it fits the points everywhere, fits better in some places, or simply mismatches the pattern. The \gD{} layout and the langevitour view with the model are linked together via rectangular brushes; when a brush is active, points will be highlighted in the adjacent view. Because the langevitour is dynamic, brush events that become active will pause the animation, so that a user can interrogate the current view. The interface is constructed as a **browsable HTML widget** specifically designed for interactive data analysis.

Brushing the high model error in the \gD{} layout, with dense C-shaped data explained in @sec-effect-dens, helps in locating the positions of the related \pD{} points in langevitour view. This process helps in identifying the high model errors that occur due to the sparse end (add the video).

@@

```{r}
#| label: fig-one-dens_clust-error
#| fig-height: 5
#| fig-width: 15
#| fig-pos: H
#| out-height: 50%
#| fig-cap: "The sparse end of the structure shows a high model error, while the dense end shows a low model error. The tSNE layout of the C-shaped structure, with dense points (a) are colored according to their high-dimensional error. Darker colors represent a high error, while lighter colors indicate a low error. b and c present the results from linked plots that highlighted the high-dimensional error points in \\gD{} and $4\\text{-}D$. Videos of the linked plots are available at <>."

>>>>

Diagnosing the model while locating points in the \gD{} layout and displaying the generated model overlaid on data in the \pD{} is important.

The \gD{} layout and the langevitour view with the model are linked together via rectangular brushes; when a brush is active, points will be highlighted in the adjacent view. Because the langevitour is dynamic, brush events that become active will pause the animation, so that a user can interrogate the current view. The interface is constructed as a **browsable HTML widget** specifically designed for interactive data analysis.

To understand how well the model fits the points—whether it fits consistently, works better in certain areas, or fails to match the overall pattern—it is important to link and brush the points with high model error in the \pD{} error plot, \gD{} layout, and the generated model overlay on the data in \pD{}.

<!-- Brushing the high model error in the \gD{} layout, with dense C-shaped data explained in @sec-effect-dens, helps in locating the positions of the related \pD{} points in langevitour view. This process helps in identifying the high model errors that occur due to the sparse end (add the video). -->

@@

```{r}
#| label: fig-one-dens_clust-error
#| fig-height: 5
#| fig-width: 15
#| fig-pos: H
#| out-height: 50%
#| fig-cap: "The sparse end of the structure shows a high model error, while the dense end shows a low model error. The tSNE layout of the C-shaped structure, with dense points (a) are colored according to their $4\\text{-}D$ error. Darker colors represent a high error, while lighter colors indicate a low error. b and c present the results from linked plots that brushed the high $4\\text{-}D$ error points in \\gD{} and $4\\text{-}D$. It helps in identifying the high $4\\text{-}D$ model errors that occur due to the sparse end. Videos of the linked plots are available at <>."
